DATA 621 Homework #4

### 2019-11-15 hw4/HW4_Mike.Rmd
Tinkering

---
title: "Mike's Scratchpad for DATA 621 Homework #4"
author: "Mike Silva"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "hide"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
require(gridExtra)
library(Amelia)
library(kableExtra)
library(caret)
library(DMwR)
library(scales)
library(RColorBrewer)
library(FFTrees)
# Thank you Stack Overflow!
# A Prefix nulling hook.

@@

## Data Cleanup

The dollar values have dollar signs.  These will be stripped out and converted from a character to numeric.  LOG10 and LN transformations of all dollar values will be preformed.  Categorical data will be translated into dummy variables.

@@

```{r}
to_dummies <- function(df, var, drop = FALSE){
  temp <- data.frame(INDEX = df$INDEX, CAT = df[[var]]) %>%
    mutate(CAT = as.character(CAT),
           VAL = 1) %>%
    spread(CAT, VAL, fill = 0)
  new_names <- c()
  for (n in names(temp)){
    if(n != "INDEX"){
      n <- paste0(var,"_",n)
      n <- gsub(" ", "_", n)
    }
    new_names <- c(new_names, n)
  }
  names(temp) <- new_names

  if(drop){
    drops <- c(var)
    df <- df[ , !(names(df) %in% drops)]
  }

  df %>%
    merge(temp)
}

get_driver_type <- function(x, y){
  if(is.na(x) | is.na(y)){
    return(NA)
  }
  if(y < 40){
    driver_type <- "Young"
  } else if (y < 70){
    driver_type <- "Middle Age "
  } else {
    driver_type <- "Older"
  }
  if( x == "M"){
    return(paste(driver_type, "Male"))
  } else {
    return(paste(driver_type, "Female"))
  }
}

temp <- df %>%
  rowwise() %>%
  mutate(INCOME = strip_dollars(INCOME),
         HOME_VAL = strip_dollars(HOME_VAL),
         BLUEBOOK = strip_dollars(BLUEBOOK),
         OLDCLAIM = strip_dollars(OLDCLAIM)) %>%
  mutate(JOB = ifelse(str_length(JOB) > 0, JOB, NA),
         CAR_AGE = ifelse(str_length(CAR_AGE) > 0, CAR_AGE, NA),
         EDUCATION = ifelse(EDUCATION == "<High_School", "LT_High_School", EDUCATION),
         URBANICITY = ifelse(URBANICITY == "Highly Urban/ Urban", "Urban", "Rural")) %>%
  mutate(DRIVER_TYPE = get_driver_type(SEX, AGE)) %>%
  ungroup() %>%
  na.omit() %>%
  to_dummies("PARENT1", T) %>%
  to_dummies("KIDSDRIV", T) %>%
  to_dummies("HOMEKIDS", T) %>%
  to_dummies("MSTATUS", T) %>%
  to_dummies("SEX", T) %>%
  to_dummies("EDUCATION", T) %>%
  to_dummies("JOB", T) %>%
  to_dummies("CAR_USE", T) %>%
  to_dummies("CAR_TYPE", T) %>%
  to_dummies("RED_CAR", T) %>%
  to_dummies("REVOKED", T) %>%
  to_dummies("MVR_PTS", T) %>%
  to_dummies("URBANICITY", T) %>%
  to_dummies("DRIVER_TYPE", T) %>%
  rowwise() %>%
  mutate(BLUEBOOK = ifelse(BLUEBOOK > 0, BLUEBOOK, 0)) %>%
  # mutate(LOG10_INCOME = log10(INCOME + 1),
  #        LN_INCOME = log(INCOME + 1),
  #        LOG10_HOME_VAL = log10(HOME_VAL + 1),
  #        LN_HOME_VAL = log(HOME_VAL + 1),
  #        LOG10_BLUEBOOK = log10(BLUEBOOK + 1),
  #        LN_BLUEBOOK = log(BLUEBOOK + 1),
  #        LOG10_OLDCLAIM = log10(OLDCLAIM + 1),
  #        LN_OLDCLAIM = log(OLDCLAIM + 1)) %>%
  mutate(PRIOR_CLAIM = ifelse(CLM_FREQ + OLDCLAIM > 0, 1,0)) %>%
  ungroup() 
```

```{r}
set.seed(42)
minority <- nrow(temp[temp$TARGET_FLAG == 1,])
majority <- nrow(temp[temp$TARGET_FLAG == 0,])
diff <- majority - minority
minority_index <- temp[temp$TARGET_FLAG == 1,]$INDEX
over_sample_df <- data.frame(INDEX = sample(minority_index, diff, T)) %>%
  merge(temp, .) %>%
  bind_rows(temp)

train_index <- createDataPartition(over_sample_df$TARGET_FLAG, p = .7, list = FALSE, times = 1)
train <- over_sample_df[train_index,]
test <- over_sample_df[-train_index,]
```

```{r}
fft_train <- train %>%
  select(-TARGET_AMT, -INDEX)#, -OLDCLAIM, -CLM_FREQ)
fft_test <- test %>%
  select(-TARGET_AMT, -INDEX)#, -OLDCLAIM, -CLM_FREQ)
fit <- FFTrees(TARGET_FLAG ~ ., fft_train, fft_test)
plot(fit)
```
The FFTree suggests that past history of filing claims is a good predictor if someone will file a claim in the future.  This could be our baseline model criteria.

Other variables that seems to come up is if the driver is in an urban location and possibly their home value and if it's a commercial vehicle

>>>>

---
title: "DATA 621 Homework #4"
author: "Critical Thinking Group 3"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "hide"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
require(gridExtra)
library(Amelia)
library(kableExtra)
library(caret)
library(DMwR)
library(scales)
library(RColorBrewer)
library(ROCR)
library(corrplot)
# Thank you Stack Overflow!
# A Prefix nulling hook.

@@

@@

## Introduction

We have been given a dataset with `r nrow(df)` records representing customers of an auto insurance company.  Each record has two response variables.

The first objective is to train a logistic regression classifier to predict if a person was in a car crash.  The response variable here is `TARGET_FLAG` respresenting whether a person had an accident (1) or did not have one (0).

```{r, echo=FALSE}
df %>% 
  mutate(TARGET_FLAG = as.factor(TARGET_FLAG)) %>%
  ggplot(aes(x=TARGET_FLAG,fill=TARGET_FLAG)) +
  geom_bar() + scale_y_continuous() + scale_fill_brewer(palette="Set1") +
  theme_light() +
  theme(legend.position = "none") +
  labs(x="TARGET_FLAG", y="# Observations")
```

The second objective will be to train a regression model to predict the cost of a crash, if one occurred.  The second reponse variable, `TARGET_VALUE`, is the amount it will cost if the person crashes their car. The value is zero if the person did not crash their car.

```{r, echo=FALSE}
df %>% filter(TARGET_FLAG == 1) %>%
  ggplot(aes(x=TARGET_AMT)) + geom_density() +
  geom_vline(aes(xintercept = mean(TARGET_AMT)), lty=2, col="red") +
  geom_label(aes(x=mean(TARGET_AMT),y=1,label="mu"),parse=T) +
  geom_vline(aes(xintercept = median(TARGET_AMT)), lty=2, col="darkgreen") +
  geom_label(aes(x=median(TARGET_AMT),y=.5,label="median")) +
  scale_x_log10(labels=comma) + theme_light() +
  labs(title="TARGET_AMT Density Plot", caption="x-axis is log 10 scale",
       y="Density", x="LOG(TARGET_AMT)")
```

Looking at the distribution of the `TARGET_AMT` variable, we can see that the variable is considerably right-skewed. Thus, a LOG transform might be best here.

## Data Exploration

We will first look at the summary statistics for the data

```{r}
df %>%
  summary() %>%
  kable() %>%
  kable_styling()
```

There are some missing values that we will need to deal with. There are also some values that seem invalid (i.e. -3 CAR_AGE).

## Data Preparation

### Fix Data Types

There are some variables that are currently factors that are dollar values that need to be transformed into numeric variables.  We will do this to both the `df` and `evaluation` data frames.  There are also some invalid data that will be changed to NAs.

```{r}
strip_dollars <- function(x){
  x <- as.character(x)
  x <- gsub(",", "", x)
  x <- gsub("\\$", "", x)
  as.numeric(x)
}

fix_data_types <- function(messy_df){
  messy_df %>%
    rowwise() %>%
    mutate(INCOME = strip_dollars(INCOME),
           HOME_VAL = strip_dollars(HOME_VAL),
           BLUEBOOK = strip_dollars(BLUEBOOK),
           OLDCLAIM = strip_dollars(OLDCLAIM)) %>%
    ungroup()
}

na_bad_values <- function(messy_df){
  messy_df %>%
    rowwise() %>%
    mutate(CAR_AGE = ifelse(CAR_AGE < 0, NA, CAR_AGE))%>%
    ungroup()
}

df$TARGET_FLAG <- factor(df$TARGET_FLAG)

df <- df %>%
  fix_data_types() %>%
  na_bad_values()
evaluation <- evaluation %>%
  fix_data_types() %>%
  na_bad_values()
```

Now that we have fixed the variables, we can look at a sumamry of the data:

```{r}
df %>%
  summary() %>%
  kable() %>%
  kable_styling()
```

### Fix Missing Values

There are `r df %>% filter(is.na(CAR_AGE)) %>% nrow()` observations where the `CAR_AGE` variable is missing, `r df %>% filter(is.na(YOJ)) %>% nrow()` observations where variable `YOJ` is missing, `r df %>% filter(is.na(AGE)) %>% nrow() ` observations where the variable `AGE` is missing, `r df %>% filter(is.na(INCOME)) %>% nrow() ` observations where the variable `INCOME` is missing, and `r df %>% filter(is.na(HOME_VAL)) %>% nrow() ` observations where the variable `HOME_VAL` is missing.  There are `r nrow(df) - nrow(na.omit(df))`, or `r round(((nrow(df) - nrow(na.omit(df))) / nrow(df)) * 100)`% of the observations missing varables.

### Feature Creation

We will create a log transformed income and home value feature.  We will also create an average claim amount that will hopefully track better with the `TARGET_AMT` variable.

```{r}
# Function to add features

create_features <- function(d){
  d %>%
    mutate(LOG_INCOME = log(INCOME + 1),
           LOG_HOME_VAL = log(HOME_VAL + 1),
           AVG_CLAIM = ifelse(CLM_FREQ > 0, OLDCLAIM / CLM_FREQ, 0),
           PRIOR_ACCIDENT = factor(ifelse(OLDCLAIM == 0 & AVG_CLAIM == 0, 0, 1)),
           COLLEGE_EDUCATED = factor(ifelse(EDUCATION %in% c("Bachelors", "Masters", "PhD"), 1, 0)),
           URBAN_DRIVER = factor(ifelse(URBANICITY == "Highly Urban/ Urban", 1, 0)),
           YOUNG_MALE = factor(ifelse(SEX == "M" & AGE < 25, 1, 0)),
           YOUNG = factor(ifelse(AGE < 25, 1, 0)),
           RED_SPORTS_CAR = factor(ifelse(CAR_TYPE == "Sports Car" & RED_CAR == "yes", 1, 0)),
           HAS_KIDS = factor(ifelse(HOMEKIDS == 0, 0, 1)),
           KID_DRIVERS = factor(ifelse(KIDSDRIV == 0, 0, 1)))
}

df <- create_features(df)
evaluation <- create_features(evaluation)
```

### Creating Training/Test Data Sets

Now that we have a complete data set we will split the data into a training (`train`) and test set (`test`). We'll use a 70-30 split between train and test, respectively.

```{r}
set.seed(42)
# ADAM THIS BROKE YOUR MODEL
#train_index <- createDataPartition(df$TARGET_AMT, p = .7, list = FALSE, times = 1)
train_index <- createDataPartition(df$TARGET_FLAG, p = .7, list = FALSE, times = 1)
train <- df[train_index,]
test <- df[-train_index,]
```

There are `r nrow(train[train$TARGET_FLAG == 1,])` out of `r nrow(train)` records in the training data set that have been in an accident.  We want to correct the imballance in the dataset by over sampling this group so our model will do a better job identifying this minority group.

```{r}
over_sample <- function(df, predictors){
  vars <- c("INDEX", "TARGET_FLAG", predictors)
  df <- df[vars] %>%
    na.omit()
  set.seed(42)
  minority <- nrow(df[df$TARGET_FLAG == 1,])
  majority <- nrow(df[df$TARGET_FLAG == 0,])
  diff <- majority - minority
  minority_index <- df[df$TARGET_FLAG == 1,]$INDEX
  over_sample_df <- data.frame(INDEX = sample(minority_index, diff, TRUE)) %>%
    merge(df, .) %>%
    bind_rows(df)
  return(over_sample_df)
}

evaluate_model <- function(model, test_df){
  test_df$yhat <- ifelse(predict.glm(model, test_df,"response") >= 0.5, 1, 0)
  cm <- confusionMatrix(factor(test_df$yhat), factor(test$TARGET_FLAG), "1")
  deviance <- model$deviance
  r2 <- 1 - model$deviance / model$null.deviance

  cat("F1 =", cm$byClass[7],"\nR2 =", r2, "\n\n")
  print(cm)
  return(cm)
}

```

## Model Creation

### Baseline Model

We will create a simple model to serve as the baseline.  This model posits that drivers history is a good representation of their future.  So drivers that have been in an accident (`OLDCLAIM` > 0 or `AVG_CLAIM` > 0) are more likely to be in an accident.  Conversely those who haven't aren't likely to be in one in the future.

```{r, null_prefix = TRUE, comment=NA}
temp <- over_sample(train, c("PRIOR_ACCIDENT"))
baseline_model <- glm(TARGET_FLAG ~ PRIOR_ACCIDENT, family = binomial(link = "logit"), temp)
summary(baseline_model)

results <- evaluate_model(baseline_model, test)
```

This simple model has statistically significant predictors.  Applying this model to the test data set indicates this simple model has a `r  round(unname(results$overall['Accuracy'])*100,1)`% accuracy rate.  It correcly identified `r round(unname(results$byClass['Sensitivity'])*100,1)`% of the people with accidents and `r  round(unname(results$byClass['Specificity'])*100,1)`% of those without.  It is better than flipping a coin but not by much.  All future models must outpreform this baseline.

### Risk Taker Model

For this model we are going assume that risk takers are more likely to be in an accident.  To identify those who tend to be risk takers, we are going with the urban legend that those who drive red sports cars are risk takers.  Also young men (16-24) also tend to be risk takers.

```{r, null_prefix = TRUE, comment=NA}
risk_taker_model <- glm(TARGET_FLAG ~ RED_SPORTS_CAR + YOUNG_MALE, family = binomial(link = "logit"), over_sample(train, c("RED_SPORTS_CAR", "YOUNG_MALE")))
summary(risk_taker_model)

risk_taker_results <- evaluate_model(risk_taker_model, test)
```

This model only has one statistically significant predictor, however it has a `r  round(unname(risk_taker_results$overall['Accuracy'])*100,1)`% accuracy rate.  This is because it identified `r  round(unname(risk_taker_results$byClass['Specificity'])*100,1)`% of those who didn't have an accident (and they are more numerous).  When looking at the model's performance on identifying the minority group of interest we see it correcly identified `r round(unname(risk_taker_results$byClass['Sensitivity'])*100,1)`% of the people with accidents.  So while the accuracy rate out preformed the baseline model, this one will not be the model we will use.

### Traditional Model

Traditional wisdom holds that there are a few tried and true predictors of someone's risk of an automobile accident. Namely: age (under 25 vs. 25+), marital status, driving record, distance to work, and the driver's sex (which may or may not be proper to use, but seems to have held throughout the years).

```{r}
traditional_model <- glm(TARGET_FLAG ~ YOUNG + MSTATUS + PRIOR_ACCIDENT + SEX +
                           REVOKED + MVR_PTS + TRAVTIME + CAR_USE,
             family = binomial(link = "logit"), over_sample(train, c("YOUNG", "MSTATUS", "PRIOR_ACCIDENT", "SEX", "REVOKED", "MVR_PTS", "TRAVTIME", "CAR_USE")))

summary(traditional_model)

model_results <- evaluate_model(traditional_model, test)
```

### Traditional Model with Cross-Validation

Here we try to use cross-validation techniques to improve the traditional model. We go back to the raw training data, because it is typically not-advised to use data that is oversampled prior to cross-validation.

```{r}
# Get complete cases
cases <- train %>%
  select(YOUNG,MSTATUS,PRIOR_ACCIDENT,SEX,REVOKED,MVR_PTS, TRAVTIME, CAR_USE) %>%
  complete.cases()

temp <- train[cases,]

# Use 5-fold cross-validation
train_control <- trainControl(method="cv",number=5, sampling="up")

traditional_cv <- train(form = TARGET_FLAG ~ YOUNG + MSTATUS + PRIOR_ACCIDENT + SEX +
                           REVOKED + MVR_PTS + TRAVTIME + CAR_USE,
                        method = "glm", family = "binomial", 
                        data = temp, trControl = train_control)

traditional_cv
```

```{r, eval=FALSE}
# Evaluating the model
eval <- data.frame(actual = test$TARGET_FLAG, predicted = predict(traditional_cv,newdata=test,type="raw"), prob = predict(traditional_cv,newdata=test,type="prob"))

confusionMatrix(eval$predicted, eval$actual, positive = "1")

pred <- prediction(eval$prob.1, eval$actual)
auc <- performance(pred, measure = "auc")@y.values[[1]]
perf <- performance(pred,"tpr","fpr")
plot(perf,main="ROC Curve", sub=paste0("AUC: ",round(auc,3)))
```

### Manual Model

```{r, null_prefix = TRUE, comment=NA}
#"KIDSDRIV"         "AGE"             
#[6] "HOMEKIDS"         "YOJ"              "INCOME"           "PARENT1"          "HOME_VAL"        
#[11] "MSTATUS"          "SEX"              "EDUCATION"        "JOB"              "TRAVTIME"        
#[16] "CAR_USE"          "BLUEBOOK"         "TIF"              "CAR_TYPE"         "RED_CAR"         
#[21] "OLDCLAIM"         "CLM_FREQ"         "REVOKED"          "MVR_PTS"          "CAR_AGE"         
#[26] "URBANICITY"       "LOG_INCOME"       "LOG_HOME_VAL"     "AVG_CLAIM"        "PRIOR_ACCIDENT"  
#[31] "COLLEGE_EDUCATED" "URBAN_DRIVER"     "YOUNG_MALE"       "RED_SPORTS_CAR"   "HAS_KIDS"        
#[36] "KID_DRIVERS"     
model <- glm(TARGET_FLAG ~ PRIOR_ACCIDENT + KID_DRIVERS + MSTATUS + INCOME +
             SEX + CAR_USE + COLLEGE_EDUCATED + REVOKED + URBAN_DRIVER,
             family = binomial(link = "logit"), over_sample(train, c("PRIOR_ACCIDENT", "KID_DRIVERS", "MSTATUS", "INCOME", "SEX", "CAR_USE", "COLLEGE_EDUCATED", "REVOKED", "URBAN_DRIVER")))
summary(model)

model_results <- evaluate_model(model, test)
```

## Amount Model

To select build a model of how much the claim amount will be we need to subset the training data to only include those that have accidents.  There will be `r train %>% filter(TARGET_FLAG == 1) %>% nrow()` data points in this set.

```{r}
amt_train <- train %>%
  filter(TARGET_FLAG == 1) %>%
  select(-INDEX, -TARGET_FLAG)
amt_test <- test %>%
  filter(TARGET_FLAG == 1)
```

```{r}
M <- amt_train %>%
  purrr::keep(is.numeric) %>%
  #mutate(TARGET_AMT = log(TARGET_AMT)) %>%
  na.omit() %>%
  cor(.)
corrplot(M, type = "upper")
```

It looks like `BLUEBOOK` is the only numeric variable that correlates with the claim amount.  We will use it as the baseline and see if we can improve this models.

```{r}
amt_fit <- lm(TARGET_AMT ~ BLUEBOOK, amt_train)
summary(amt_fit)
```
